Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-21 latex/article.tex
wip section II article.tex

\sloppy\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widely investigated since the
seminal Marching-Cubes approach by Lorensen and
Cline~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by Lengyel \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs.  However, it
was only developed on CPU to visualize in real-time static or near
static data, combining octree with new Marching-Cubes configurations
to generate adaptive meshes. In this paper we propose a GPU pipeline
inspired from this method.  It leaves the CPU free for other tasks and
avoids synchronization with the GPU.  Our high framerate also allows
this method to run in real-time on fully dynamic data.  The
triangulation is controlled by a GPU LoD criterion that ensures that
triangles project onto more than one pixel, thus removing geometrical
aliasing.

@@

\sloppy In our context, we consider digital shapes (any subset $Z$ of
$\Z^d$) and boundaries of digital shapes $\partial Z$. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid
step $h$, \emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset
of $\R^d$ corresponding to the union of hypercubes centered at $h\cdot
\vz$ for $\vz\in Z$ with edge length $h$. By doing so, both
$\partial\Shape$ and $\partial\Body{Z}{h}$ are topological boundary of
objects lying in the same space (See Fig. \ref{fig:notations}-b). Note
that the combinatorial digital boundary of $Z$ made of cells in a
cellular Cartesian complex (\emph{pointels}, \emph{linels}, {\em
  surfels}, \ldots), can be trivally embedded into $\R^d$ such that it
coincides with $\partial\Body{Z}{h}$.

@@

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $V_{R_d}$ (or $A_{R_d}$ in dimension 2) of the
intersection between the shape $\Shape$ and a ball of digital radius $R_d$:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature estimators} are defined for any point $\vx
  \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes, where $k$ is a constant
related to the maximal curvature of the shape, with at least $C^3$-boundary and
bounded curvature) \cite{DGCI2013}. For short, when the digital objet becomes
finer and finer, \ie when the digitization step $h$ tends to zero, the estimated
quantity on $\Bd{\Body{\DSh}{h}}$  converges (theoretically and experimentally)
to the associated one on $\Bd{\Shape}$. More precisely, we have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

In \cite{CVIU2014}, we have also defined the 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $\Body{\DSh}{h}$ of a shape $\Shape \subset \R^3$
  scaled by the digitization step $h$, {\em 3D digital principal curvature
  estimators} are defined for any point $\vx \in \R^3$ as:
  %
  \begin{align}
    \PrincCurvH{1}{R}(\Body{\DSh}{h},\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5{{R}}} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(\Body{\DSh}{h},\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5{{R}}} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues of
  the covariance matrix of the digital points in the intersection between
  $\Body{\DSh}{h}$ and the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on
  $\vx/h$.\todo[inline]{J: Pareil ici... pas top...}
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$, $1$ and $2$
(see Equation~20 of \cite{CVIU2014} for more details). We have also proved the
convergence of these estimators when setting the ball radius $h$ in $R =
kh^\frac{1}{3}$, where $k$ is a constant related to the maximal curvature of the
shape, on convex shapes with at least $C^3$-boundary and bounded curvature
\cite{CVIU2014}. More formally, we get:
%
\begin{align}
  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
\end{align}

>>>>

\sloppy\noindent\textbf{Related works}\quad Extracting and Visualizing
isosurface on volumetric data has been widely investigated since the
seminal Marching-Cubes approach by Lorensen and
Cline~\cite{lorensen1987marching}. This approach being data parallel,
GPU implementation of this method is very efficient
\cite{tatarchuk2007real}. However, such technique generates a lot of
triangles which is not well suited to large digital data. Hence,
adaptive approaches have been proposed in order to optimize the
triangular mesh resolution to the object geometry or depending on the
viewpoint. In this topic, many solutions have been developed in
Computer
Graphics~\cite{shu1995adaptive,schaefer2004dual,lengyel2010voxel,DBLP:journals/cgf/LewinerMPPL10,DBLP:journals/cvgip/LobelloDD14}.
The method developed by Lengyel \textit{et al.}
\cite{lengyel2010voxel} is the best suited to our needs.  However, it
was developed to visualize in real-time static or near static data and
present a complex CPU implementation.\todo{DC: rajouter une phrase par
  rapport à lengyel, Helene ? JP?}  In this paper we propose a GPU
pipeline inspired from this method.  It leaves the CPU free for other
tasks and avoids synchronization with the GPU.  Our high framerate
also allows this method to run in real-time on fully dynamic data.
The triangulation is controlled by a GPU LoD criterion that ensures
that triangles project onto more than one pixel, thus removing
geometrical aliasing.

@@

\sloppy In our context, we consider digital shapes (any subset $Z$ of $\Z^d$)
and boundaries of digital shapes $\DigBoundary{Z}$. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset of $\R^d$
corresponding to the union of hypercubes centered at $h\cdot \vz$ for $\vz\in Z$
with edge length $h$. By doing so, both $\partial\Shape$ and
$\partial\Body{Z}{h}$ are topological boundary of objects lying in the same
space (See Fig. \ref{fig:notations}-b). Note that the combinatorial digital
boundary $\DigBoundary{Z}$ of $Z$ made of cells in a cellular Cartesian complex
(\emph{pointels}, \emph{linels}, \emph{surfels}, \ldots), can be trivally
embedded into $\R^d$ such that it coincides with $\partial\Body{Z}{h}$.

@@

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ with
$\MCard(Y)$ the number of digital points of $Y$ and $Y \subset \Z^d$ the
intersection between the digital shape $\DigShape$ and a ball of digital radius
$R/h$ (we denote such quantity $\AreaC(Y,h)$ in dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property \cite{DGCI2013}. For short, when the digital objet becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
associated one on $\Bd{\Shape}$.
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
Setting $R = kh^\frac{1}{3}$ on convex shapes
with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
related to the maximal curvature of the shape, we have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

In \cite{CVIU2014}, we have also defined the 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z \EqDef \DSh$ of a shape $\Shape \subset \R^3$,
  {\em 3D digital principal curvature estimators} are defined for any point $\vx
  \in \R^3$ as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    \PrincCurvH{1}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5 R} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5 R} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues
  of the covariance matrix of the intersection between the digital shape
  $\DigShape$ and a ball of digital radius $R/h$ centered on digital point
  $\vx/h$.
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$
($\DMom{000}{h}$), of order $1$ ($\DMom{100}{h}$, $\DMom{010}{h}$ and
$\DMom{001}{h}$) and of order $2$ ($\DMom{200}{h}$, $\DMom{020}{h}$,
$\DMom{002}{h}$, $\DMom{110}{h}$, $\DMom{011}{h}$ and $\DMom{101}{h}$, see
Equation~20 of \cite{CVIU2014} for more details). We have also proved the
convergence of these estimators when setting the ball radius $h$ in $R =
kh^\frac{1}{3}$, where $k$ is a constant related to the maximal curvature of the
shape, on convex shapes with at least $C^3$-boundary and bounded curvature
\cite{CVIU2014}. More formally, we get:
%
\begin{align}
  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
\end{align}
